---
title: "Learning and Forecasting of Age–Specific Period Mortality via B–Spline Processes with Locally–Adaptive Dynamic Coefficients"
author: "Federico Pavone, Sirio Legramanti, and Daniele Durante"
date: "`r format(Sys.time(), '%d/%m/%Y')`"
output:
  html_document:
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE,
                      class.source = "foldable",
                      fig.align = 'center')
```

The purpose of this notebook is to show step-by-step how to implement and fit the proposed BSP model. Refer to the paper for a detailed explanation of the model. The complete code to generate all the results of the paper can be run via `main.R`.

As an example, we consider Italian mortality from 1933 to 2019. 

# Working Environment Setup

If you have not done it before, run the following line to create the `Rdata` file containg the mortality rates for Italy and other countries.
```{r data_processing}
library(here)
source(here('source','data_processing.R'))
```

Then, load and source all the necessary libraries and functions. The package `KFAS` is used to implement the state-space formulation of the model, whereas the other packages are mostly to handle data and to generate plots. We source also functions in the `source` folder.
```{r setup_environment}
library(tidyverse)
library(KFAS)
library(doParallel)
library(plotly)
library(ggh4x)
library(ggrepel)
library(patchwork)
library(latex2exp)
library(reactable)
source(here('source','BSP.R'))
source(here('source','setup.R'))
source(here('source','helper_fun.R'))
set.seed(4238)
theme_set(
  theme_light() +
    theme(
      strip.background = element_rect(color = 'gray', fill = 'white'),
      strip.text.x = element_text(color = 'black'),
      strip.text.y = element_text(color = 'black')
    )
)
```

# Data Loading and Exploration

Let's load the data and visualize it. We create a 3d-plot of the log-mortality rates of Italy (overall population), highlighting years 1970 and 2005, and ages 23 and 60. 

We can notice the overall decreasing trend of the mortality across all ages, together with generally higher mortality for newborns and older people. Moreover, we see how the age-wise behaviour is smoother than the year-wise one. The shock in the '40s is due to World War II.
```{r data, class.source = 'fold-hide'}
load(here('output','mortality.Rdata'))

as_tibble(log((Y_ita_man + Y_ita_woman)/(N_ita_man + N_ita_woman))) %>%
  mutate(year = years[1:(length(years)-1)]) %>%
  pivot_longer(-year, names_to = 'age', values_to = 'value') %>%
  mutate_at(vars(age), as.numeric) %>%
  pivot_wider(names_from = age, values_from = value) %>%
  select(-year) %>%
  as.matrix() %>% 
  t() -> data3d_ita
colnames(data3d_ita) <- years[1:(length(years)-1)]

plot_ly(z=~data3d_ita,
        x = years[1:(length(years)-1)]) %>%
  add_surface(alpha = 0.8,
              colors = "Spectral",
              reversescale = TRUE) %>%
  add_trace(x = rep(1970, 101), 
            y = 0:100, 
            z = data3d_ita[,'1970'], 
            mode = 'lines',
            type = 'scatter3d',
            line = list(color = '#037A73',
                        width = 6)) %>%
  add_trace(x = rep(2005, 101), 
            y = 0:100, 
            z = data3d_ita[,'2005'], 
            mode = 'lines',
            type = 'scatter3d',
            line = list(color = '#037A73',
                        width = 6)) %>%
  add_trace(x = years[1:(length(years)-1)],
            y = 23, 
            z = data3d_ita[24,], 
            mode = 'lines',
            type = 'scatter3d',
            line = list(color = '#7570b3',
                        width = 6)) %>%
  add_trace(x = years[1:(length(years)-1)],
            y = 60, 
            z = data3d_ita[61,], 
            mode = 'lines',
            type = 'scatter3d',
            line = list(color = '#7570b3',
                        width = 6)) %>%
  layout(scene = list(xaxis = list(title = 'Years', showgrid = F,
                                   linecolor = 'black',
                                   linewidth = 1.5),
                      yaxis = list(title = 'Age', showgrid = F,
                                   linecolor = 'black',
                                   linewdith = 1.5),
                      zaxis = list(title = 'log(Rate)', showgrid = F,
                                   linecolor = 'black',
                                   linewidth = 1.5),
                      camera = list(
                        center = list(x = 0, y = 0, z = -0.1),
                        eye = list(
                          x = 1.5,
                          y = 0.9,
                          z = 1.5))),
         showlegend = FALSE) %>%
  hide_colorbar()

```

We create a list `data_list` with separated rates for males and females to be used in the rest of the analysis.
```{r}
data_list <- list(ita_man = Y_ita_man/N_ita_man,
                  ita_woman = Y_ita_woman/N_ita_woman)
```

# Model Definition and Fit

We define our model according to the setting specified in `source/setup.R`. In particular, the spline basis is the following:
```{r spline, class.source = 'fold-hide'}
S <- mSpline(ages[-1], knots = age_knots, degree = 2, intercept = TRUE)
K <- ncol(S)
for (k in 1:K){
  S[,k] <- S[,k]/max(S[,k])
}
colnames(S) <- paste('U', 1:K, sep = '')
ages_max <- c('U0'=0, apply(S, 2, which.max))

colors_spline <- c("black" = 'grey80',
                   "color" = '#2c7fb8')
colors_spline_names <- ifelse(paste('U',0:K,sep='') %in% c('U3','U8','U15'), 
                              '#2c7fb8', 
                              'black')

Sfull <- bSpline(seq(1, 100, by = 0.1), knots = age_knots, degree = 2, intercept = TRUE)
for (k in 1:K){
  Sfull[,k] <- Sfull[,k]/max(Sfull[,k])
}
colnames(Sfull) <- paste('U', 1:K, sep = '')

ages_max_Sfull <- seq(1,100,by=0.1)[apply(Sfull,2,which.max)]


UtoG_Tex <- paste('$g_{',20:1,'}(x)$',sep='') %>%
  sapply(X = .,
         FUN = . %>% latex2exp::TeX(input = .), USE.NAMES = FALSE)
  
cbind('U0' = rep(0, nrow(S)), S) %>%
  rbind(c(1, rep(0, ncol(S))),.) %>%
  as_tibble() %>%
  mutate(age = ages) %>%
  pivot_longer(-age, names_to = 'weight', values_to = 'value') %>%
  mutate_at(vars(value), as.numeric) %>%
  mutate(label = ifelse(weight %in% c('U3','U8','U15'), "color", 'black')) %>%
  ggplot(aes(x = age,
             y = factor(weight, 
                        levels = paste('U', K:0, sep = '')))) +
  geom_tile(aes(fill = label, alpha = value), colour = "white", size = 0.4) +
  labs(y = NULL, x = NULL) +
  guides(fill = "none", alpha = "none") +
  scale_x_continuous(breaks = seq(0,100,by=5), limits = c(-0.5,100)) + #, sec.axis = dup_axis()) +
  scale_y_discrete(position = 'left', labels = UtoG_Tex) +
  scale_fill_manual(values = colors_spline) +
  scale_alpha_continuous(range = c(0,1)) +
  theme(panel.grid.minor.y = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        text=element_text(size = 9),
        plot.margin = margin(t = 0),
        axis.text.y = element_text(colour = rev(colors_spline_names)),
        axis.ticks.y = element_blank(),
        axis.ticks.x = element_blank()) -> plot_spline

as_tibble(Sfull)  %>%
  mutate(age = seq(1, 100, by = 0.1)) %>%
  pivot_longer(-age, names_to = 'weight', values_to = 'value') %>%
  mutate_at(vars(value), as.numeric) %>%
  mutate(label = ifelse(weight %in% c('U3','U8','U15'), "color", 'black')) %>%
  ggplot(aes(x = age, y = value)) +
  geom_line(aes(group = weight, color = label, size = label), alpha = 0.8) +
  geom_segment(aes(x = 0, y = 0, xend = 0, yend = 1),
              size = 0.3,
              color = colors_spline['black']) +
  geom_point(data = tibble(x = c(0,ages_max_Sfull), y = 1), aes(x=x, y=y),
             shape = 5,
             size = 0.9) +
  geom_label_repel(data = tibble(x = c(0,ages_max_Sfull),
                           y = 1,
                           label = ages_max),
             aes(x = x, y = y, label = label),
             size = 2,
             segment.size = 0.2,
             segment.alpha = 0.5,
             label.size = 0.1,
             ylim = c(1.05,NA),
             min.segment.length = 0.1) +
  scale_size_manual(values = c("black" = 0.4, "color" = 0.8)) +
  scale_color_manual(values = colors_spline) +
  scale_x_continuous(limits = c(-0.5,100)) +
  scale_y_continuous(limits = c(0,1.5)) +
  labs(x = NULL, y = NULL) +
  guides(color = "none", alpha = "none", size = "none") +
  geom_segment(aes(x = 0, y = 0, xend = 100, yend = 0), 
               color = "grey90", size = 1) +
  theme(panel.grid = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.margin = margin(b = 0),
        axis.ticks.y = element_blank()) -> plot_spline2

plot_spline2 / plot_spline + 
  plot_layout(height = c(1,2))
```

The model definition is obtained via the `bsp.model` function. Since we are working on two datasets simultaneously, all operations are done list-wise using `lapply`. 

**Note**: for those who wants to inspect the implementation of `bsp.model` or other functions in `source`, they should keep in mind that the notation of the code uses letter $U$ referring to parameters $\beta$ in the paper.
```{r model}
model_list <- lapply(data_list, 
                     . %>% bsp.model(rates = .,
                                      delta = delta,
                                      age_knots = age_knots,
                                      kernel = matern_kernel))
```

Once the `KFAS` model is defined, we estimate the hyperparameters using the `bsp.fit` function. To guarantee robustness of the result, it is advisable to repeat the optimization with different starting points. The function allows to specify the argument `rep` and `parallel` to possibly parallelize the optimization with `rep` different starting points. In `source/setup.R`, the default is `rep = 100`. Moreover, following Zhu and Dunson (2013) our implementation includes a penalization on the parameters $\sigma_\beta$ and $\sigma_a$ equivalent to the Inverse-Gamma prior used in their work. Check function `bsp.fit` in `source/BSP.R` for further details.
```{r fit, eval = FALSE}
fit_list <- lapply(model_list,
                   . %>%
                     bsp.fit(., 
                              rep = rep, 
                              method = 'Nelder-Mead', 
                              parallel = TRUE, 
                              maxcl = 50))
```

If parallelization is not feasible for the user, the above code might take long time to run. For this purpouse, we provide the result of the fit in the file `output/ITA_fit.Rdata` (alternatively, one can run the above chunk with a smaller number of `rep`). 

```{r load_fit}
# If previous chunk not executed
load(here('output','ITA_fit.Rdata'))
#
```

The estimated parameters are
```{r table, class.source = 'fold-hide'}
fit_list %>%
  modify(. %>% 
           pluck('info') %>%
           pluck('optim') %>%
           pluck('par') %>%
           exp()) %>%
  imap_dfr(.f = ~ .) %>%
  mutate(par = c('lambda','sigma2_u','sigma2_a','sigma2_e')) %>%
  pivot_longer(-par, names_to = 'data', values_to = 'value') %>%
  pivot_wider(names_from = 'par', values_from = 'value') %>%
  reactable(columns = list(
    data = colDef(name = 'Data'),
    lambda = colDef(name = "Lambda", format = colFormat(digits = 3)),
    sigma2_u = colDef(name = "Sigma2_beta", format = colFormat(digits = 3)),
    sigma2_a = colDef(name = "Sigma2_a", format = colFormat(digits = 3)),
    sigma2_e = colDef(name = "Sigma2_phi", format = colFormat(digits = 3))
  ))
```
Notice that the values of `sigma2_beta`, `sigma2_a`, `sigma2_e` should be related to the scale-parameter `lambda` before comparisons across different data and fits.

# Posterior Distribution via Kalman Smoothing

It is possible to obtain posterior means and variances of the latent states $\beta_{t}$s via Kalman smoothing using the `KFS` function of the `KFAS` library. Specifying the argument `filtering` in `KFS`, it is also possible to obtain the filtering distribution (see documentation of `KFS`).
```{r smoothing}
smooth_list <- lapply(fit_list, . %>% 
                        pluck('fit') %>%
                        KFS(., smoothing = c('mean', 'state', 'signal'),
                            simplify = FALSE, 
                            maxiter = 200))
```

Once obtained the smoothing distribution, we can prepare plot friendly datasets using the functions available in `source/helper_fun.R` (e.g. `smoothing2tibble`), and plot the results.

```{r}
lambda_est <- fit_list %>%
  modify(. %>%
           pluck('info') %>%
           pluck('optim') %>%
           pluck('par') %>%
           exp() %>%
           `[`(.,1)) %>%
  imap_dfr(.f = ~ .)

smoothing_tibble <- lapply(smooth_list, . %>% 
                                       smoothing2tibble(., country = 'ITA')) %>%
                                bind_rows(.id = 'data') %>%
                                mutate(gender = sub(".*_", "", data),
                                       country = sub("_.*", "", data)) %>%
                                mutate(state_gender = paste(state, gender, sep = '_'))

smoothing_tibble_adj <- smoothing_tibble %>%
  filter(state %in% c('U','dU')) %>%
  mutate(lambda_delta = case_when(state == 'U' ~ 1,
                                  data == 'ita_man' ~ lambda_est$ita_man*delta,
                                  data == 'ita_woman' ~ lambda_est$ita_woman*delta)) %>%
  mutate_at(vars(value, value.sd), ~ .*lambda_delta) %>%
  mutate_at(vars(gender), ~ ifelse(. == 'man', 'Male', 'Female')) %>%
  mutate_at(vars(country), .f = toupper)
```

For example, we can plot the smoothing distribution of $\beta(t)$ together with the first derivative $\partial\beta(t)/\partial t$.
```{r, class.source = 'fold-hide'}
age_classes <- tibble(weight = 'U0',
                      range = 'Newborn')
for(i in 1:K){
  range0 <- which(S[,i] > 0.2)
  age_classes <- bind_rows(age_classes,
                           tibble(weight = paste('U', i, sep = ''),
                                  range = paste(min(range0), max(range0), sep = '~')))
}
age_classes <- age_classes %>%
  mutate(max = ages_max) %>%
  mutate(spline = paste('S',0:K,sep=''))

smoothing_tibble_adj %>%
  left_join(age_classes, by = "weight") %>%
  filter(weight %in% c('U0','U7','U11'),
         gender == 'Female') %>%
  mutate(label_dem = case_when(range == 'Newborn' ~ 'Infant (age 0)',
                               range == '23~32' ~ 'Young (age 23 ~ 32)',
                               range == '46~64' ~ 'Adult (age 46 ~ 64)')) %>%
  mutate_at(vars(label_dem), . %>% factor(.,
                                          levels = c('Infant (age 0)',
                                                     'Young (age 23 ~ 32)',
                                                     'Adult (age 46 ~ 64)'))) %>%
  ggplot(aes(x = t, y = value)) + 
  geom_segment(data = tibble(x = 1933,
                             y = 0,
                             xend = 2020,
                             yend = 0,
                             state = 'dU'),
               aes(x = x, y = y, xend = xend, yend = yend),
               alpha = 0.5, linetype = 'dashed') +
  geom_line(color = '#abdda4',
            #alpha = 0.6,
            size = 0.6) +
  geom_ribbon(aes(ymin = value - 2*value.sd,
                  ymax = value + 2*value.sd),  
              fill = '#abdda4',
              alpha = 0.4) +
  facet_nested(factor(state, levels = c('U', 'dU'),
                      labels = c(U = latex2exp::TeX(r'($\beta$)'),
                                 dU = latex2exp::TeX(r'(d$\beta$)'))) ~  gender + label_dem,
               scale = 'free',
               labeller = labeller(.rows = label_parsed),
               strip = strip_nested(
                 text_x = elem_list_text(colour = c("black", "black")),
                 background_x = elem_list_rect(fill = c("grey95", "white")),
                 by_layer_x = TRUE)) +
  ggh4x::force_panelsizes(rows = c(2, 1)) + 
  facetted_pos_scales(y = list(scale_y_continuous(n.breaks = 3), 
                               scale_y_continuous(n.breaks = 3, limits = c(-0.5,0.5)))) +
  scale_x_continuous(breaks = c(1940,1980,2020)) +
  labs(linetype = NULL,y = NULL,x = NULL) +
  theme(legend.position = "none",
        strip.text.y.right = element_text(angle = 0),
        text = element_text(size = 9),
        plot.margin = margin(b = 0),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_line(linetype = 'dashed')) -> plot_female

smoothing_tibble_adj %>%
  left_join(age_classes, by = "weight") %>%
  filter(weight %in% c('U0','U7','U11'),
         gender == 'Male') %>%
  mutate(label_dem = case_when(range == 'Newborn' ~ 'Infant (age 0)',
                               range == '23~32' ~ 'Young (age 23 ~ 32)',
                               range == '46~64' ~ 'Adult (age 46 ~ 64)')) %>%
  mutate_at(vars(label_dem), . %>% factor(.,
                                          levels = c('Infant (age 0)',
                                                     'Young (age 23 ~ 32)',
                                                     'Adult (age 46 ~ 64)'))) %>%
  ggplot(aes(x = t, y = value)) + 
  geom_segment(data = tibble(x = 1933,
                             y = 0,
                             xend = 2020,
                             yend = 0,
                             state = 'dU'),
               aes(x = x, y = y, xend = xend, yend = yend),
               # alpha = 0.5,
               linetype = 'dashed') +
  geom_line(color = "#abdda4",
            alpha = 0.6,
            size = 0.6) +
  geom_ribbon(aes(ymin = value-2*value.sd, 
                  ymax = value+2*value.sd),
              fill = "#abdda4",  alpha = 0.4) +
  facet_nested(factor(state, levels = c('U', 'dU'),
                      labels = c(U = latex2exp::TeX(r'($\beta$)'),
                                 dU = latex2exp::TeX(r'(d$\beta$)'))) ~  gender + label_dem,
               scale = 'free',
               labeller = labeller(.rows = label_parsed),
               strip = strip_nested(
                 text_x = elem_list_text(colour = c("black", "black")),
                 background_x = elem_list_rect(fill = c("grey95", "white")),
                 by_layer_x = TRUE)) +
  force_panelsizes(rows = c(2, 1)) + 
  facetted_pos_scales(y = list(scale_y_continuous(n.breaks = 3), 
                               scale_y_continuous(n.breaks = 3, limits = c(-0.5,0.5)))) +
  scale_x_continuous(breaks = c(1940,1980,2020)) +
  labs(linetype = NULL,y = NULL,x = NULL) +
  theme(legend.position = "none",
        strip.text.y.right = element_text(angle = 0),
        text = element_text(size = 9),
        plot.margin = margin(t = 0),
        axis.title.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_line(linetype = 'dashed')) -> plot_male


plot_female / plot_male
```

# Male Mortality Forecast via Random Walk plus Drift

As described in the paper, mortality forecasts are obtained via a **random walk with drift** model estimated over the smoothing means of the last 25 years. The code to fit and forecast $h$ step ahead is availble in `source/BSP_forecast.R` through the function `fitandforecast`. This function first call `bsp.fit` to fit the model, and then build the forecast via the random walk plus drift model. 

Since we have already fit the model, we extract and use here only the part of `fitandforecast` that defines, fits, and forecasts the random walk plus drift model.

For the sake of simplicity of the illustration in this notebook, we restrict the forecast step only to one of the two datasets (e.g. male mortality). Clearly, the procedure is general, and applicable to any dataset (also simply via `lapply` with the `fitandforecast` function).

The random walk plus drift model is defined as (notation of the paper):
$$
\beta_{t_{s^*+1}} = \beta_{t_{s^*}} + \hat{\lambda}\delta_{s^*}\hat{\Delta} + \omega_{t_{s^*}} \qquad \omega_{t_{s^*}}\overset{\text{i.i.d.}}{\sim}N_p(0,W),
$$
with $W_{[j,l]}=\sigma^2_\omega\rho_{\beta_{[j,l]}}$.

The following code computes $\hat{\Delta}$ as the median of the smoothing distribution of $\beta_{t}$ over the last 25 years, and estimates $\sigma^2_\omega$ via maximum likelihood on the mean of the smoothing distribution of $\beta_{t}$. The function `optim_rwd` in `source/BSP_forecast.R` provides the likelhood function of the random walk plus drift model.

```{r rw_fit}
source(here('source','BSP_forecast.R'))
fit <- fit_list$ita_man
smooth <- smooth_list$ita_man
train <- length(years)
n_for <- 25
h <- 10
# The following code is extracted and adapted from the fitandforecast() function
K <- fit$info$K
Z <- fit$info$Z
## Fitting random-walk with drif for prediction
# Extracting info on U and dU from BSP model
U_m <- smooth$alphahat[, paste('U', 0:K, sep = '')][train - n_for + 1,]
U <- smooth$alphahat[, paste('U', 0:K, sep = '')] %>% 
  tail(n = n_for, k = 1) # selecting last n_for years
dU <- apply(smooth$alphahat[, paste('dU', 0:K, sep = '')] %>%
              tail(n = n_for, k = 1), # selecting last n_for years
            MARGIN = 2, median)
## Creating random walk + dirft model
Tt_pred <- fit$fit$T[1,2,1] # lambda*delta
kernel <- fit$info$kernel
ages_max <- fit$info$ages_max
# Estimation of rw + drift
rep_pred <- 5
starting_values <- log(runif(rep_pred, min = 1e-2, max = 2))
fit_rw_list <- lapply(starting_values, 
                      function(init){
                        try(optim(par = init,
                                  fn = optim_rwd,
                                  method = "L-BFGS-B",
                                  control = list(maxit = 1e6),
                                  u = U, 
                                  mu0 = U_m,
                                  Tt = Tt_pred,
                                  drift = dU,
                                  kernel = fit$info$kernel,
                                  ages_max = fit$info$ages_max,
                                  K = K), TRUE)
                      })
# Checking failed optim
fit_rw_list_clean <- discard(fit_rw_list, 
                             . %>% inherits(., 'try-error'))
if(length(fit_rw_list_clean) == 0){
  print('All optimization attempts failed..')
}
# Extracting best fit
best_fit_rw <- which.min(sapply(fit_rw_list_clean, . %>% `$`(.,value)))
fit_rw <- fit_rw_list_clean[[best_fit_rw]]
```

Once the random walk with drift model is estimated, we can forecast $\beta_t$ for, e.g., the next 10 years. The forecast for the age-specific mortality rates can be then straightforwardly obtained by linear combination with the spline basis functions.
```{r forecast}
## Composing rw + drift with estimated sigma2_u
sigma2_u <- exp(fit_rw$par)
Qt_pred <- diag(K+1) * sigma2_u
for(I in 1:(K+1-1)){
  for(J in (I+1):(K+1)){
    rho_u <- kernel(x = abs(ages_max[J]-ages_max[I]))
    Qt_pred[I,J] <- Qt_pred[J,I] <- sigma2_u * rho_u
  }
}
Zt_pred <- fit$fit$Z[, seq(1, 3*(K+1), by = 3), ]
a1_pred <- tail(U, n = 1 ,k = 1) # last year of training
P1_pred <- smooth$V %>%
  tail(n = c(3*(K+1), 3*(K+1), 1)) %>% # selecting last year 
  # extracting only var(u) components
  `[`(seq(1, 3*(K+1), by = 3), seq(1, 3*(K+1), by = 3), ) 
# Prediction h-step ahead
U_pred <- matrix(NA, nrow = h, ncol = K+1)
f_pred <- matrix(NA, nrow = h, ncol = Z)
for(tt in 1:h){
  if(tt == 1){
    U_pred[1,] <- as.numeric(a1_pred) + Tt_pred%*%dU
    f_pred[1,] <- Zt_pred%*%U_pred[1,]
  } else {
    U_pred[tt,] <- U_pred[tt-1,] + Tt_pred%*%dU
    f_pred[tt,] <- Zt_pred%*%U_pred[tt,]
  }
}

Upred <- list()
for(kk in 1:(K+1)){
  Upred <- c(Upred,
            list(tibble(fit = U_pred[,kk])))
}

fpred <- list()
for(zz in 1:Z){
  fpred <- c(fpred,
             list(tibble(fit = f_pred[,zz])))
}
```

For the purpose of the illustration, we report here the obtained forecasts for the next 10 years of the spline weights $\hat{\beta}_{t^*j}$, for $j=1,\dots,20$.
```{r beta_forecast}
Upred %>% 
  imap_dfr(.f = ~ (as_tibble(.) %>%
             mutate(h_ahead = 1:h)),
           .id = 'weight') %>%
  mutate_at(vars(weight), as.numeric) %>%
  mutate(t = tail(years,1) - 1 + h_ahead) -> forecast_beta

forecast_beta %>%
  mutate(label = paste('Beta', weight, sep = '')) %>%
  select(label, fit, t) %>%
  mutate_at(vars(fit), . %>% round(.,digits = 2)) %>%
  pivot_wider(names_from = 'label', values_from = 'fit') %>%
  reactable(columns = list(t = colDef(name = 'Year')))
```
Finally, we plot the forecasted age-specific mortality curves.
```{r plot_forecast}
fpred %>% 
  imap_dfr(.f = ~ (as_tibble(.) %>%
             mutate(h_ahead = 1:h)),
           .id = 'age') %>%
  mutate_at(vars(age), as.numeric) %>%
  mutate(t = tail(years,1) - 1 + h_ahead) -> forecast_age

forecast_age %>%
  ggplot(aes(x = age, y = fit, color = factor(t))) +
  geom_line() +
  geom_label_repel(data = . %>% 
                     filter(t == 2020,
                            age == 30),
                   aes(x = age, y = fit, label = t),
                   nudge_y = 0.9,
                   nudge_x = 1) +
  geom_label_repel(data = . %>% 
                     filter(t == 2029,
                            age == 15),
                   aes(x = age, y = fit, label = t),
                   nudge_y = -0.5,
                   nudge_x = 15) +
  scale_color_brewer(palette = 'BrBG') +
  scale_y_continuous(breaks = c(-1.5,-4.5,-7.5)) +
  scale_x_continuous(n.breaks = 10) +
  labs(x = "Age") +
  guides(color = "none") +
  theme(panel.grid = element_blank(),
        axis.title.y = element_blank())
```






